\section{Model Description}
\label{sec:model_description}

Models descriptions are formatted in XML (Extensible Markup Language) tag
structures to allow easy human and computer readability, and allow easier collaborations between
developers writing applications that interact with model definitions.

The model XML document has a structure that is defined by a schema.
The schema of the XML document is currently located
at:

%\begin{mylisting}
\begin{verbatim}
  http://flame.ac.uk/schema/xmml_v3.xsd
\end{verbatim}
%\end{mylisting}

This provides a way to validate the model document to make sure all the tags are
being used correctly. This can be achived by using xml command line tools like
XMLStarlet \cite{xmlstarlet} and xmllint \cite{xmllint} or by using editors that
can have xml validation builtin like Eclipse \cite{eclipse}. The start and end
of a model file should be formatted as follows:

%\begin{mylisting}
\begin{verbatim}
  <?xml version="1.0" encoding="ISO-8859-1"?>
  <!DOCTYPE xmodel SYSTEM "http://flame.ac.uk/schema/xmml_v3.xsd">
  <xmodel version="3">
  <name>Model_name</name>
  <version>the version</version>
  <description>a description</description>
  ...
  </xmodel>
\end{verbatim}
%\end{mylisting}

Where name is the name of the model, version is the version, and description
allows the description of the model. Models can also contain:

\begin{itemize}
\item \textbf{other models} (enabled or disabled)
\item \textbf{environment}
\begin{itemize}
\item constant variables
\item location of function files
\item time units
% \begin{itemize}
% \item name
% \item *** unit
% \item *** period
% \end{itemize}
\item data types
% \begin{itemize}
% \item name
% \item description
% \item variables
% \end{itemize}
\end{itemize}
\item \textbf{agent types}
\begin{itemize}
\item name
\item description
\item memory
% *** variables
\item functions
% *** name
% *** description
% *** current state
% *** next state
% *** condition
% *** inputs
% **** filter
% *** outputs
\end{itemize}
\item \textbf{message types}
\begin{itemize}
\item name
\item description
\item variables
\end{itemize}
\end{itemize}

\subsection{Model in Multiple Files}

It is possible to define a model in multiple files. FLAME reads a model from
multiple files as if the model was defined in one file. This capability allows
different parts of a model to be enabled or disabled easily. For example if a
model includes different versions of a sub-model that can be exchanged, or a
subsystem of a model can be disabled to see how it affects the model.
Alternatively this capability could be used as a hierarchy, for example a `body'
model could include a model of the `cardiovascular system' that includes a
model of the `heart'. The following tags show the inclusion of two models, one
enabled and one disabled:

%\begin{mylisting}
\begin{verbatim}
  <models>
    <model><file>sub_model_1.xml</file><enabled>true</enabled></model>
    <model><file>sub_model_2.xml</file><enabled>false</enabled></model>
  </models>
\end{verbatim}
%\end{mylisting}

The location of any sub model should be relative to the original model xmml
file.

\subsection{Environment}

The environment of a model holds information that maybe required by a model but
is not part of an agent or a message. This includes:

\begin{itemize}
\item constant variables -- for setting up different simulations easily
\item location of function files -- the path to the implementations of agent
functions
\item time units -- for easily activating agent functions dependent on time
periods
\item data types -- user defined data types used by agent memory or
message variables
\end{itemize}

This notion of environment does not correspond to an environment that would be
a part of a model where agents would interact with the environment. Anything
that can change in a model must be represented by an agent, therefore if a
model includes a changeable environment that agents can interact with, this in
itself must be represented by an agent.

\subsubsection{Constant Variables}

These are constant variables that can be set at the start of simulation runs
initial starting values and cannot be altered thereafter. They can be defined as
follows:

%\begin{mylisting}
\begin{verbatim}
  <constants>
    <variable>
     <type>int</type>
     <name>my_constant</name>
     <description>value read in initial simulation settings</description>
    </variable>
  </constants>
\end{verbatim}
%\end{mylisting}

Varible names cannot have spaces, instead please use lower case letters,
numbers, underscores and hyphens. This also applies to agent memory names, message names, data type
names and function names.

% Constant Variables refers to the global values used in the model. These can me
% defined in a separate header file which can then be included in one of the
% functions file.
% 
% The header file would look as follows:
% 
% \begin{mylisting}
% \begin{verbatim}
% #define <varname> <value>
% \end{verbatim}
% \end{mylisting}
% 
% If this file was saved as a `my\_header.h' file, include this file into one of
% the function files so that the compiler knows about these arguments.

\subsubsection{Function Files}

Function files hold the source code for the implementation of the
agent functions. The location is relative to the location of the model
xmml file. They are included in the compilation script (Makefile) of the
produced model:

%\begin{mylisting}
\begin{verbatim}
  <functionFiles>
  <file>function_source_code_1.c</file>
  <file>function_source_code_2.c</file>
  </functionFiles>
\end{verbatim}
%\end{mylisting}

\subsubsection{Time Units}
\label{timeunit}

% Time units allow the possibility of restricting the functions to
% only execute during particular iterations.
% Time rules can be applied to function conditions instead of a
% condition rule and are defined by a time period and a phase. A time
% phase is the offset from the start of a period.

Time units are used to define time periods that agent functions act within. For
example a model that uses a calendar based time system could take a day to be
the smallest time step, i.e. one iteration. Other time units can then use this
definition to define other time units, for example weeks, months, and years.

A time unit contains:

\begin{itemize}
\item name -- name of the time unit.
\item unit -- can contain `iteration' or other defined time units.
\item period -- the length of the time unit using the above units.
\end{itemize}

An example of a calendar based time unit set up can be defined as:

%\begin{mylisting}
\begin{verbatim}
  <timeUnits>
  
    <timeUnit>
      <name>daily</name>
      <unit>iteration</unit>
      <period>1</period>
    </timeUnit>
    
    <timeUnit>
      <name>weekly</name>
      <unit>daily</unit>
      <period>5</period>
    </timeUnit>
    
    <timeUnit>
      <name>monthly</name>
      <unit>weekly</unit>
      <period>4</period>
    </timeUnit>
    
    <timeUnit>
      <name>quarterly</name>
      <unit>monthly</unit>
      <period>3</period>
    </timeUnit>
    
    <timeUnit>
      <name>yearly</name>
      <unit>monthly</unit>
      <period>12</period>
    </timeUnit>
    
  </timeUnits>
\end{verbatim}
%\end{mylisting}

\clearpage

\subsubsection{Data Types}

Data types are user defined data types that can be used in a model. They are a
structure for holding variables. Single variables can be:

\begin{itemize}
  \item C fundamental data types -- int, float, double, char.
  \item static array -- of size ten: variable\_name[10].
  \item dynamic array -- available by placing `\_array' after
  the data type name: variable\_name\_array (please do not use if possible as
  useage affects load balancing on high performance computers).
\end{itemize}

For example:

\begin{verbatim}
  <variables>
    <variable>
      <type>int</type>
      <name>int_single</name>
      <description>A single integer</description>
    </variable>
    <variable>
      <type>int</type>
      <name>int_list[2]</name>
      <description>A list of 2 integers</description>
    </variable>
    <variable>
      <type>int_array</type>
      <name>int_dynamic_list</name>
      <description>A list of integers that can change size</description>
    </variable>
  </variables>
\end{verbatim}

Data types can hold the above types of variables and also other defined data
types. In the example below the data type \emph{line} contains a variable
of data type \emph{position} which is defined above it:

%\begin{mylisting}
\begin{verbatim}
  <dataTypes>

 <dataType>
  <name>position</name>
  <description>position in 3D using doubles</description>
  <variables>
   <variable><type>double</type><name>x</name>
    <description>position on x-axis</description>
   </variable>
   <variable><type>double</type><name>y</name>
    <description>position on y-axis</description>
   </variable>
   <variable><type>double</type><name>z</name>
    <description>position on z-axis</description>
   </variable>
  </variables>
 </dataType>

 <dataType>
  <name>line</name>
  <description>a line defined by two points</description>
  <variables>
   <variable><type>position</type><name>start</name>
    <description>start position of the line</description>
   </variable>
   <variable><type>position</type><name>end</name>
    <description>end position of the line</description>
   </variable>
  </variables>
 </dataType>

  </dataTypes>
\end{verbatim}
%\end{mylisting}

\subsection{Agents}

An agent type contains a name, a description, memory, and functions:

%\begin{mylisting}
\begin{verbatim}
  <agents>

  <xagent>
    <name>Agent_Name</name>
    <description></description>
    <memory>
     ...
    </memory>
    <functions>
      ...
    </functions>
  </xagent>
  
  </agents>
\end{verbatim}
%\end{mylisting}
% 
%   <xagent>
%     <name>Household</name>
%     <description></description>
%     <memory>
%       <variable><type>int</type><name>id</name>
%        <description></description>
%       </variable>
%       <variable><type>int</type><name>region_id</name>
%        <description></description>
%       </variable>
%       <variable><type>int_array</type><name>neighboring_region_ids</name>
%        <description></description>
%       </variable>
%       <variable><type>int</type><name>gov_id</name>
%        <description></description>
%       </variable>
%       <variable><type>int</type><name>day_of_month_to_act</name>
%        <description></description>
%       </variable>
%       <variable><type>double</type><name>payment_account</name>
%        <description></description>
%       </variable>
%     </memory>
%     <functions>
%      <function>
%        <name>Household_read_firing_messages</name>
%         <description>Check for being fired or not</description>
%         <currentState>EXIT_FINANCIAL_MARKET</currentState>
%         <nextState>01d</nextState>
%         <condition>
%          <lhs><value>a.employee_firm_id</value></lhs>
%          <op>NEQ</op>
%          <rhs><value>-1</value></rhs>
%         </condition>
%         <inputs>
%          <input><messageName>firing</messageName></input>
%         </inputs>
%       </function>
%     </functions>
%   </xagent>
% </agents>


\subsubsection{Agent Memory}

Agent memory defines variables, where variables are defined by their type, C
data types or user defined data types from the environment, a name, and a
description:

%\begin{mylisting}
\begin{verbatim}
  <memory>
    <variable>
      <type>int</type>
      <name>id</name>
      <description>identity number</description>
    </variable>
    <variable>
      <type>double</type>
      <name>x</name>
      <description>position in x-axis</description>
    </variable>
  </memory>
\end{verbatim}
%\end{mylisting}

Note: Agent memory variable cannot be called 'name' as this is used to define the agent type in simulation input files.

\subsubsection{Agent Functions}

An agent function contains:

\begin{itemize}
\item name - the function name which must correspond to an implemented function
name and must be unique across the model
\item description
\item current state - the current state the agent has to be in.
\item next state - the next state the agent will transition to.
\item condition - a possible condition of the function transition.
\item inputs - the possible input messages.
\item outputs - the possible output messages.
\end{itemize}

And as tags:

%\begin{mylisting}
\begin{verbatim}
  <function>
    <name>function_name</name>
    <description>function description</description>
    <currentState>current_state</currentState>
    <nextState>next_state</nextState>
    <condition>
    ...
    </condition>
    <inputs>
    ...
    </inputs>
    <outputs>
    ...
    </outputs>
  </function>
\end{verbatim}
%\end{mylisting}

The current state and next state tags hold the names of states. This is the
only place where states are defined. State names must coordinate with other
functions states to produce a transitional graph from a single start state to
end many possible end states.

%\paragraph{Function Condition}
\label{functioncond}

A function can have a condition on its transition. This condition can include
conditions on the agent memory and also on any time units defined in the
environment. At any state with outgoing transitions with conditions it must be
possible for a transition to happen, i.e. it must be possible for every agent
to transition from the start state to an end state. Each possible transition
must be mutually exclusive, i.e. the order that the function conditions are
tested is not defined. A function named `idle' is available to be used for
functions that do not require an implementation.

Conditions (that are not just time unit based) take the form:

\begin{itemize}
  \item lhs -- left hand side of comparison
  \item op -- the comparison operator
  \item rhs -- the right hand side of the comparison
\end{itemize}

Or in tags:

%\begin{mylisting}
\begin{verbatim}
  <lhs></lhs><op></op><rhs></rhs>
\end{verbatim}
%\end{mylisting}

Sides to compare (lhs or rhs) can be either a value, denoted by value tags, 
a formula, currently also in value tags, or another comparison rule.
Values and formula can include agent variables which are preceded by `a'.

% \begin{mylisting}
% \begin{verbatim}
% a.agent_var
% m.message_var
% \end{verbatim}
% \end{mylisting}

The comparison operator can be one of the following comparison functions:

\begin{itemize}
\item EQ -- equal to
\item NEQ -- not equal to
\item LEQ -- less than or equal to
\item GEQ -- greater than or equal to
\item LT -- less then
\item GT -- greater than
\item IN -- an integer (in lhs) is a member of an array of integers (in rhs)
\end{itemize}

or can be one of the following logic operators as well:

\begin{itemize}
\item AND
\item OR
\end{itemize}

The operator `NOT' can be used by placing `not' tags around a comparison rule.
For example the following tagged rule describes the condition being true when
the `z' variable of the agent is greater than zero and less than ten:

%\begin{mylisting}
\begin{verbatim}
  <condition>
    <lhs>
      <lhs><value>a.z</value></lhs>
      <op>GT</op>
      <rhs><value>0.0</value></rhs>
    </lhs>
    <op>AND</op>
    <rhs>
      <not>
        <lhs><value>a.z</value></lhs>
        <op>LT</op>
        <rhs><value>10.0</value></rhs>
      </not>
    </rhs>
  </condition>
\end{verbatim}
%\end{mylisting}

A condition can also depend on any time units described in the environment. For
example the following condition is true when the agent variable
`day\_of\_month\_to\_act' is equal to the number of iterations since of the
start, the phase, of the `monthly' period, i.e. twenty iterations as defined in
the time unit:

%\begin{mylisting}
\begin{verbatim}
  <condition>
    <time>
      <period>monthly</period>
      <phase>a.day_of_month_to_act</phase>
    </time>
  </condition>
\end{verbatim}
%\end{mylisting}

Time conditions can also be encased within a `not' tag in indicate the reverse
condition. \\

% The condition allows the function to run \emph{monthly} at the phase
% of \emph{day\_of\_month\_to\_act}. The
% \emph{day\_of\_month\_to\_act} is a variable extracted from the
% agent memory and is thus defined as
% \emph{a.day\_of\_month\_to\_act}.
% 
% Refer to section \ref{functioncond} for more details on function
% condition definitions.
% 
% These rules are then parsed into rule functions and placed in a file
% called rules.c

Functions can have input and output message types. For example the following
example the function takes message types `a' and `b' as inputs and outputs
message type `c':

%\begin{mylisting}
\begin{verbatim}
  <inputs>
    <input><messageName>a</messageName></input>
    <input><messageName>b</messageName></input>
  </inputs>
  <outputs>
    <output><messageName>c</messageName></output>
  </outputs>
\end{verbatim}
%\end{mylisting}

%\paragraph{Message Filter}
Message filters can be applied to message inputs and allow the messages to be
filtered. Filters are defined similar to function conditions but include
message variables which are prefixed by an `m'. The following filter only
allows messages where the agent variable `id' is equal to the message variable
`worker\_id':

%\begin{mylisting}
\begin{verbatim}
  <input>
    <messageName>firing</messageName>
    <filter>
      <lhs><value>a.id</value></lhs>
      <op>EQ</op>
      <rhs><value>m.worker_id</value></rhs>
    </filter>
    <sort>
      <key>wage</key>
      <order>descend</order>
    </sort>
  </input>
\end{verbatim}
%\end{mylisting}

The previous example also includes the use of a sort tag. The sort tag uses a
key, a message variable, to sort the list in `ascend'ing or `descend'ing order
via the order tag.

Using filters in the model description enables FLAME to make message
communication more efficient by pre-sorting messages and using other techniques.

% Thus in the above example messages will be filtered according to the
% message variable \emph{worker\_id} (defined as m.<varname>) to be EQ
% (equal) to the agent \emph{id} (defined as a.<varname>).

\subsection{Messages}

Messages defined in a model must have a type which is defined by a name and the
variables that are included in the message. The following example is a message
called `signal' that holds a position in 3D.
\clearpage
%\begin{mylisting}
\begin{verbatim}
  <messages>
  
  <message>
    <name>signal</name>
    <description>Holds the position of the sending agent</description>
    <variables>
      <variable>
        <type>double</type>
        <name>x</name>
        <description>The x-axis position</description>
      </variable>
      <variable>
        <type>double</type>
        <name>y</name>
        <description>The y-axis position</description>
      </variable>
      <variable>
        <type>double</type>
        <name>z</name>
        <description>The z-axis position</description>
      </variable>
    </variables>
  </message>
  
  </messages>
\end{verbatim}
%\end{mylisting}
